---
title: "cellbrowser"
author: "Kent Riemondy RBI"
date: "2/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cellbrowser

Format seurat object into a cellbrowser directory suitable for html browser. 

```{r}
library(tidyverse)
atcells <- readRDS("preprocess/s_obj_final.rds")
```

### Metadata

```{r add_mdata}

mdata <- atcells@meta.data
new_mdata <- tibble::rownames_to_column(mdata, "cell") %>% 
  dplyr::mutate(sample_type = ifelse(str_detect(new_expt_id, 
                                                 "Non-ATII epithelial"),
                                            "Naive Non-AEC2 Epithelial",
                                            ifelse(str_detect(new_expt_id,
                                                              "ATII expt"),
                                                   "Naive AEC2",
                                                   "Injured AEC2")),
               experiment = ifelse(expt == "expt1", 
                                   "Experiment 1",
                                   "Experiment 2")) %>% 
  dplyr::select(cell, sample_type, experiment) %>% 
  tibble::column_to_rownames("cell")

atcells <- AddMetaData(atcells, new_mdata)


atcells <- AddMetaData(atcells, 
                       data.frame(row.names = rownames(atcells@meta.data),
                                  technical_replicate = str_match(atcells@meta.data$sample_names, "([0-9])expt")[, 2]))
                        
atcells <- AddMetaData(atcells,
                       data.frame(row.names = rownames(atcells@meta.data),
                                  experiment_and_tech_rep = str_c(atcells@meta.data$experiment,
                                  ": Technical Replicate ",
                                  atcells@meta.data$technical_replicate)))

```

Add pretty labels to cell annotations

```{r pretty_labels}
cell_id_relabel <- c("Endothelial" = "Endothelial/Fibroblast",
  "Basal" = "Basal",
  "Club" = "Club",
  "Ciliary" = "Ciliated",
  "Naive Type I" = "Naive AEC1",
  "Macrophages" = "Macrophage",
  "Injured Type II" = "Other Injured AEC2",
  "Proliferating Type II" = "Injured AEC2: Proliferating",
  "Cell Cycle Arrest Type II" = "Injured AEC2: Cell Cycle Arrest",
  "Transdifferentiating Type II" = "Injured AEC2: Transdifferentiating",
  "Naive Type II" = "Naive AEC2")


m_df <- data.frame(row.names = rownames(atcells@meta.data),
                       pretty_cell_labels = cell_id_relabel[atcells@meta.data$labeled_clusters])

m_df$pretty_cell_labels <- factor(m_df$pretty_cell_labels,
                                  levels = cell_id_relabel)

atcells <- AddMetaData(atcells, m_df)
```


Simplify object metadata to make interpretable

```{r}

cols <- colnames(atcells@meta.data)

cols_to_keep <- c(
  `genes per cell` = "nGene",
  `UMIs per cell` = "nUMI",
  `condition` = "sample_type",
  `technical_replicate` = "technical_replicate",
  `GFP expression` = "gfp_counts", 
  `proporition of UMIs as mitochondria` = "proportion.mito",
  `clusters` = "res.0.6",
  `cell_types` = "pretty_cell_labels")


atcells@meta.data <- atcells@meta.data[, cols_to_keep]

colnames(atcells@meta.data) <- names(cols_to_keep)
```



```{r}

outdir <- "cellbrowser"
dir.create(outdir)

saveRDS(atcells, "cellbrowser/sobj.rds")
```

Build cellbrowser objects. Note requires UCSC (`cellbrowser python package`[ https://cellbrowser.readthedocs.io/index.html])

Use bash to run this for now, but might be able to use `reticulate` R package to communicate with the python package.

```{bash}
cd cellbrowser
/miniconda3/bin/cbImportSeurat2 \
  -i sobj.rds \
  -o Lung_Regeneration_LPS \
  -m \
  -c "cell_types" 

```


```{bash}
cd cellbrowser

/miniconda3/bin/cbBuild \
 -i Lung_Regeneration_LPS/cellbrowser.conf \
 -o lung

```