---
title: "Figures"
author: "Kent Riemondy RBI"
date: '`r Sys.Date()`'
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    fig_caption: yes
    fig_retina: 1 
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs, message=FALSE, warning=FALSE, echo=FALSE}
source("../../R/globals.R")
data_dir <- file.path(data_dir, "updated")
```

## Plotting Functions

### tsne


```{r tsne_plot}

#' set up defaults for tsne plot
plot_tsne <- function(seurat_obj,
                      ident = "ident",
                      pt.size = 0.5,
                      pt.alpha = 1,
                      label_text = FALSE,
                      label.size = 6,
                      label.color = "grey",
                      .cols = NULL,
                      legend_names = NULL,
                      cell_filter = NULL,
                      legend_order = NULL,
                      ...){

  
  tsne_dat <- seurat_obj@dr$tsne@cell.embeddings %>% 
    as.data.frame() %>% 
    tibble::rownames_to_column("cell")
    
  xlimits <- c(min(tsne_dat$tSNE_1),
               max(tsne_dat$tSNE_1))
  ylimits <- c(min(tsne_dat$tSNE_2),
               max(tsne_dat$tSNE_2))
  
  if (!is.null(cell_filter)){
    seurat_obj <- SubsetData(seurat_obj, cells.use = cell_filter)
  }
  
  mdata <- seurat_obj@meta.data %>% tibble::rownames_to_column("cell")
  tsne_dat <- left_join(mdata, tsne_dat, by = "cell")

  p <- ggplot(tsne_dat, 
         aes(tSNE_1, tSNE_2)) +
    ylim(ylimits) +
    xlim(xlimits)
    
  if (is.null(ident)){
    if (is.null(.cols)){
      p <- p + geom_point(
               size = pt.size,
               alpha = pt.alpha) 
    } else {
      p <- p + geom_point(
               color = .cols[1],
               size = pt.size,
               alpha = pt.alpha) 
    }
  } else {
    p <- p + geom_point(aes_string(color = ident),
               size = pt.size,
               alpha = pt.alpha)
  }
  
  p <- p + guides(colour = guide_legend(override.aes = list(size = 4))) +
    theme(legend.title = element_blank())
  
  if (label_text) {
    tsne_mean_dat <- tsne_dat %>% 
      group_by_at(vars(one_of(ident))) %>% 
      summarize(mean_dim_1 = mean(tSNE_1), 
             mean_dim_2 = mean(tSNE_2))
    
    p <- p + 
      geom_text(data = tsne_mean_dat, 
                aes_string(x = "mean_dim_1",
                           y = "mean_dim_2",
                           label = ident),
                size = label.size, color = label.color)
  }

  ## handle colors
  if (is.null(.cols)){
      if (is.null(legend_names)){
        p <- p + scale_color_viridis(discrete = T,
                        direction = -1)
      } else {
        p <- p + scale_color_viridis(labels = legend_names,
                                     discrete = T,
                        direction = -1)
      }
  } else {
      if (is.null(legend_names) & is.null(legend_order)){
         p <- p + scale_color_manual(values = .cols)
      } else if(is.null(legend_order)) {
         p <- p + scale_color_manual(labels = legend_names, values = .cols)
      } else {
        if(is.null(legend_names)) legend_names <- legend_order
        p <- p + scale_color_manual(values = .cols, 
                                    breaks = legend_order,
                                    labels = legend_names)
      }
  }
  p
}


```

### feature

```{r feature}
#' set up defaults for tsne plot
plot_feature <- function(seurat_obj,
                      ident = NULL,
                      gene = NULL,
                      plot_dat = NULL,
                      plot_col = NULL,
                      meta_data_col = NULL,
                      pt.size = 0.001,
                      pt.alpha = 0.25,
                      label_text = FALSE,
                      label.size = 6,
                      .cols = NULL,
                      legend_names = NULL,
                      cell_filter = NULL,
                      palette_type = "cloupe",
                      col_pal = "Reds",
                      max_y = NULL,
                      legend_title = NULL,
                      ...){
  
  mdata <- seurat_obj@meta.data %>% tibble::rownames_to_column("cell")
  
  tsne_dat <- seurat_obj@dr$tsne@cell.embeddings %>% 
    as.data.frame() %>% 
    tibble::rownames_to_column("cell")
  
  tsne_dat <- left_join(mdata, tsne_dat, by = "cell")
  
  if (!is.null(cell_filter)){
    tsne_dat <- dplyr::filter(tsne_dat,
                                cell %in% cell_filter)
  }
  
  if (!is.null(gene) & is.null(meta_data_col)) {
    gene_dat <- FetchData(seurat_obj, gene) %>% 
      as.data.frame() %>% 
      tibble::rownames_to_column("cell")
    tsne_dat <- left_join(tsne_dat, gene_dat, by = "cell")
  }
  
  if (!is.null(plot_dat)){
    tsne_dat <- left_join(tsne_dat, plot_dat, by = "cell")
  }
  
  if (!is.null(ident)){
    color_aes_str <- ident
  } else if (!is.null(plot_dat)){
    color_aes_str <- plot_col
  } else if (!is.null(meta_data_col)){
    color_aes_str <- gene
  } else {
    color_aes_str <- gene
  }
  
  color_aes_str_q <- quo(color_aes_str)
  tsne_dat <- tsne_dat %>% arrange_at(.vars = color_aes_str)
  
  p <- ggplot(tsne_dat, 
         aes(tSNE_1, tSNE_2)) +
    geom_point(aes_string(color = color_aes_str),
               size = pt.size,
               alpha = pt.alpha)
  
  ## handle legend limit 
  
  if (is.null(max_y) & is.null(ident)) {
    max_y <- c(0, max(tsne_dat[[color_aes_str]]))
  } else if (!is.null(ident) & is.null(max_y)){
    max_y <- c(NA, NA)
  } 
  
  
  # loupe-like colors
  cols <- rev(brewer.pal(11, "RdGy")[c(1:5, 7)])
  
  #handle legend name
  if(is.null(legend_title)) legend_title <- color_aes_str
  ## handle zero expression
  
  if (all(max_y == c(0, 0))){
    p <- p + scale_color_gradient(low = cols[1], high = cols[1], name = legend_title)
    return(p)
  }
  
  ## handle colors
  if (is.null(.cols)){
    if (palette_type == "viridis") {
        p <- p + scale_color_viridis(discrete = F,
                        direction = -1,
                        option = col_pal,
                        limits = max_y, name = legend_title)
    } else if (palette_type == "brewer") {
        p <- p + scale_color_distiller(limits = max_y,
                                       palette = col_pal,
                                       direction = 1, name = legend_title)
    } else if (palette_type == "cloupe") {
      
        p <- p + scale_color_gradientn(limits = max_y,
                                       colors = cols, name = legend_title)
    }
  } 
  p
}
```

### violin 

```{r violin}
#' set up defaults for violin plot
plot_violin <- function(df, .x, .y, 
                        .fill = "",
                        .size = 0.50,
                        .width = 1, 
                        .scale = "width",
                        .alpha = 1,
                        cols = scale_fill_viridis(discrete = T),
                        single_col = NULL,
                        jitter = F){
  
  p <- ggplot(df, aes_string(x = .x, y = .y))
  
  if (jitter){
    p <- p  + geom_jitter(size = 0.1, alpha = 0.2, color = "black")
  }
  
  if (!is.null(single_col)){
    p <- p + 
      geom_violin(size = .size,
                  scale = .scale,
                  fill = single_col,
                  alpha = .alpha)
  } else {
    p <- p +
      geom_violin(aes_string(fill = .fill), 
                size = .size,
                scale = .scale,
                alpha = .alpha) +
    cols 
  }
  

  p <- p + theme(axis.text.x = element_text(angle = 90, 
                                     hjust = 1, 
                                     vjust = 0.5),
          legend.title = element_blank()) 
  p
}


#' set up defaults for violin plot
plot_box <- function(df, .x, .y, 
                        .fill = "",
                        .size = 0.50,
                        .width = 1, 
                        ...){
  
  ggplot(df, aes_string(x = .x, y = .y)) +
    geom_boxplot(aes_string(fill = .fill), 
                 coef = 1e6
                 ) +
    scale_fill_viridis(discrete = T) +
    theme(axis.text.x = element_text(angle = 90, 
                                     hjust = 1, 
                                     vjust = 0.5),
          legend.title = element_blank()) 
}

```


## Figure 1 code 
  
Load in original seurat object with proper filtered data. 

```{r load_dat}
atcells <- readRDS("s_obj2.rds")
atcells <- SetAllIdent(atcells, "res.1.6")
```

```{r write_out_expression}

if (!file.exists("count_matrix/expr_matrix.tsv.gz")) {
  mat <- as.data.frame(as.matrix(atcells@data))
  mat <- tibble::rownames_to_column(mat, "gene")
  write_tsv(mat, "count_matrix/expr_matrix.tsv") 
  R.utils::gzip("count_matrix/expr_matrix.tsv", 
                overwrite = T)
  rm(mat)
}
```

```{r add_mdata}

mdata <- atcells@meta.data
new_mdata <- tibble::rownames_to_column(mdata, "cell") %>% 
  dplyr::mutate(sample_type = ifelse(str_detect(new_expt_id, 
                                                 "Non-ATII epithelial"),
                                            "Naive Non-AT-II Epithelial",
                                            ifelse(str_detect(new_expt_id,
                                                              "ATII expt"),
                                                   "Naive AT-II",
                                                   "Injured AT-II")),
               experiment = ifelse(expt == "expt1", 
                                   "Experiment 1",
                                   "Experiment 2")) %>% 
  dplyr::select(cell, sample_type, experiment) %>% 
  tibble::column_to_rownames("cell")

atcells <- AddMetaData(atcells, new_mdata)

```
### Process unfiltered data
 
  Loading and filtering code largely reproduced from [filtering RMarkdown](preprocess.Rmd). From now on the replicate cell samples (i.e. ATI1expt1 and ATI2expt1) will be presented as 1 sample (AT1expt1).


```{r fig1_simple_summaries}
dir.create(file.path("fig1", "unfiltered"), recursive = TRUE, showWarnings = FALSE)
## simple summaries
merge_samples <- tibble::tribble(
  ~expt_id, ~new_expt_id,
  "ATI1expt1", "Non-ATII epithelial expt.1",
  "ATI1expt2", "Non-ATII epithelial expt.2",
  "ATI2expt1", "Non-ATII epithelial expt.1",
  "ATI2expt2", "Non-ATII epithelial expt.2",
  "ATII1expt1", "ATII expt.1", 
  "ATII1expt2", "ATII expt.2", 
  "ATIIinjured1expt1", "ATII-injured expt.1", 
  "ATIIinjured1expt2", "ATII-injured expt.2", 
  "ATIIinjured2expt1", "ATII-injured expt.1", 
  "ATIIinjured2expt2", "ATII-injured expt.2"
)

summary_dat <- data_frame(cell = colnames(atcells@raw.data),
                          expt_id = str_split(colnames(atcells@raw.data),
                                                    "_", simplify = T) %>% .[, 1], 
                          nUMIs = colSums(as.matrix(atcells@raw.data)),
                          nGenes = colSums(as.matrix(atcells@raw.data) > 0))
summary_dat <- left_join(summary_dat, 
                         merge_samples, by = "expt_id")
summary_dat <- summary_dat %>% 
  group_by(new_expt_id) %>% 
  dplyr::arrange(desc(nUMIs))

n_cells <- summary_dat %>%  
  group_by(new_expt_id) %>%  
  summarize(n_cells = n(),
            median_UMIs = median(nUMIs),
            median_Genes = median(nGenes))

n_cell_plt <- ggplot(n_cells, 
            aes(new_expt_id, n_cells)) +
  geom_bar(aes(fill = new_expt_id), stat = "identity") + 
  scale_fill_viridis(discrete = T) +
  labs(y = "Number of Cells") +
  theme(legend.pos = "none",
        axis.text.x = element_text(angle = 90, 
                                     hjust = 1, 
                                     vjust = 0.5),
        axis.title.x = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10))

n_umi_plt <- plot_violin(summary_dat, 
            "new_expt_id", "nUMIs", 
            .fill = "new_expt_id") + 
  labs(y = "Number of UMIs") +
  theme(axis.title.x = element_blank(),
        legend.pos = "none",
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10))

n_gene_plot <- plot_violin(summary_dat, 
            "new_expt_id", "nGenes", 
            .fill = "new_expt_id") + 
  labs(y = "Number of Genes\n(at least 1 UMI)") +
  theme(axis.title.x = element_blank(),
        legend.pos = "none",
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10))

save_plot(file.path("fig1", "unfiltered", "numis.pdf"), n_umi_plt,
          base_height = 4, 
          base_aspect_ratio = .9)

save_plot(file.path("fig1", "unfiltered", "ngenes.pdf"), n_gene_plot,
          base_height = 4, 
          base_aspect_ratio = .9)

save_plot(file.path("fig1", "unfiltered", "ncells.pdf"), n_cell_plt, 
          base_height = 4, 
          base_aspect_ratio = .9)
```

```{r n_mito}
mito.genes <- grep("^mt-", rownames(atcells@raw.data), value = T)
proportion.mito <- Matrix::colSums(atcells@raw.data[mito.genes, ]) /
  Matrix::colSums(atcells@raw.data)

unfiltered <- data_frame(cell = colnames(atcells@raw.data),
                         proportion.mito) %>% 
  left_join(summary_dat, by = c("cell")) %>% 
  as_data_frame()


n_mito_plot <- plot_violin(unfiltered, 
            "new_expt_id", "proportion.mito", 
            .fill = "new_expt_id") + 
  labs(y = "Proportion of reads\nderived from mitochondria") +
  theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.pos = "none")

save_plot(file.path("fig1", "unfiltered", "nmito.pdf"), n_mito_plot, 
          base_height = 4, 
          base_aspect_ratio = .9)
```

```{r n_mito_vs_umis}

n_mito_umi_plot <- ggplot(unfiltered, 
       aes(nUMIs, proportion.mito)) + 
  geom_point(aes(color = new_expt_id), 
             size = 0.25) +
  labs(x = "UMIs per cell",
       y = "Proportion of reads\nderived from mitochondria") +
  scale_color_viridis(discrete = TRUE, alpha = 0.33, name = "") +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  theme(legend.pos = c(0.05, 0.80),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 8))

save_plot(file.path("fig1", "unfiltered", "nmito_umi.pdf"), n_mito_umi_plot, 
          base_height = 4, 
          base_aspect_ratio = 1.1)

n_mito_umi_plot_zoom <- ggplot(unfiltered, 
       aes(nUMIs, proportion.mito)) + 
  geom_point(aes(color = new_expt_id), 
             size = 0.25) +
  coord_cartesian(xlim = c(0, 45000), ylim = c(0, 0.05)) + 
  labs(x = "UMIs per cell",
       y = "Proportion of reads\nderived from mitochondria") +
  scale_color_viridis(discrete = TRUE, alpha = 0.33, name = "") +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  theme(legend.pos = "top",
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 8))

save_plot(file.path("fig1", 
                    "unfiltered", 
                    "zoom_nmito_umi.pdf"), n_mito_umi_plot_zoom, 
          base_height = 4, 
          base_aspect_ratio = 1.1)
```

```{r gfp_counts_2, message = F, eval = T}

expt1_files <- dir(file.path(data_dir, "20170623_zeemans"),
    pattern = "*_sorted_counts.txt", full.names = T, recursive = T)
expt2_files <- dir(file.path(data_dir, "20170728_zeeman_expt2"),
    pattern = "*_sorted_counts.txt", full.names = T, recursive = T)


files <- c(expt1_files, expt2_files)

samples <- c(
  "3161-ATII-1",
  "3161-ATII-2",
  "3161-ATII-3",
  "3162-ATII-5",
  "3162-ATTII-4")

sample_paths_exp1 <- file.path(data_dir,
                               "20170623_zeemans", 
                               samples,
                               "outs", "filtered_gene_bc_matrices", "mm10")

expt_names <- c("ATI1expt1", "ATI2expt1", 
                "ATII1expt1", "ATIIinjured2expt1", "ATIIinjured1expt1",
                "ATIIinjured1expt2", "ATIIinjured2expt2", 
                "ATI1expt2", "ATI2expt2", "ATII1expt2")

dat <- map(files, read_tsv, col_names = c("cells", "gfp_counts"))

dat <- map2(dat, expt_names, ~dplyr::mutate(.x, group = .y))

dat <- bind_rows(dat) 

dat <- dplyr::mutate(dat, 
              cells = str_split(cells, "-", simplify = T)[, 1],
              cells = paste( group, cells, sep = "_" )) %>% 
  dplyr::select(-group)

# only keep filtered cells
dat <- dat %>% dplyr::filter(cells %in% atcells@cell.names)
# add in not-detected cells
dat <- left_join(data_frame(cells = atcells@cell.names), 
                 dat, by = c("cells"))

dat[is.na(dat)] <- 0
dat <- as.data.frame(dat)

rownames(dat) <- dat[, 1]
dat[, 1] <- NULL

# log Normalize
dat <- as.data.frame(as.matrix(LogNormalize(dat) ))


# extract out for plotting
unfiltered <- dat %>% 
  tibble::rownames_to_column("cell") %>% 
  left_join(summary_dat, by = c("cell")) %>% 
  as_data_frame()

n_gfp_plot <- plot_violin(unfiltered, 
            "new_expt_id", "gfp_counts", 
            .fill = "new_expt_id", jitter = F) + 
  labs(y = "GFP expression") +
  theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.pos = "none")

save_plot(file.path("fig1", "unfiltered", "ngfp.pdf"), n_gfp_plot, 
           base_height = 3, 
          base_aspect_ratio = .9)
```


```{r single_page_fig}
plts <- list(
  n_cell_plt,
  n_gene_plot,
  n_umi_plt,
  n_mito_plot,
  n_gfp_plot,
  n_mito_umi_plot
)

plt_top <- plot_grid(plotlist = plts[1:6],
          nrow = 3,
          ncol = 3,
         # align = 'hv',
          axis = "tl")


plt <- plot_grid(plt_top, plts[[6]], 
                 nrow = 2, ncol = 1,
                 rel_heights = c(1.5, 1))

save_plot(file.path("fig1", "unfiltered", "unfiltered_combined.pdf"), 
          plt_top, base_width = 8, base_height = 11)
```

### Process filtered data

```{r}
atcells <- readRDS("s_obj2.rds")
atcells <- SetAllIdent(atcells, "res.1.6")
```


```{r fig1_other_summaries}
dir.create(file.path("fig1", "filtered"), recursive = TRUE, showWarnings = FALSE)
## simple summaries

summary_dat <- atcells@meta.data 

n_cell_plt <- ggplot(n_cells, 
            aes(new_expt_id, n_cells)) +
  geom_bar(aes(fill = new_expt_id), stat = "identity") + 
  scale_fill_viridis(discrete = T) +
  labs(y = "Number of Cells") +
  theme(legend.pos = "none",
        axis.text.x = element_text(angle = 90, 
                                     hjust = 1, 
                                     vjust = 0.5),
        axis.title.x = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10))

n_umi_plt <- plot_violin(summary_dat, 
            "new_expt_id", "nUMI", 
            .fill = "new_expt_id") + 
  labs(y = "Number of UMIs") +
  theme(axis.title.x = element_blank(),
        legend.pos = "none",
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10))

n_gene_plot <- plot_violin(summary_dat, 
            "new_expt_id", "nGene", 
            .fill = "new_expt_id") + 
  labs(y = "Number of Genes\n(at least 1 UMI)") +
  theme(axis.title.x = element_blank(),
        legend.pos = "none",
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10))

save_plot(file.path("fig1", "filtered", "numis.pdf"), n_umi_plt,
          base_height = 4, 
          base_aspect_ratio = .9)

save_plot(file.path("fig1", "filtered", "ngenes.pdf"), n_gene_plot,
          base_height = 4, 
          base_aspect_ratio = .9)

save_plot(file.path("fig1", "filtered", "ncells.pdf"), n_cell_plt, 
          base_height = 4, 
          base_aspect_ratio = .9)
```

```{r n_mito_prop}


n_mito_plot <- plot_violin(summary_dat, 
            "new_expt_id", "proportion.mito", 
            .fill = "new_expt_id") + 
  labs(y = "Proportion of reads\nderived from mitochondria") +
  theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.pos = "none")

save_plot(file.path("fig1", "filtered", "nmito.pdf"), n_mito_plot, 
          base_height = 4, 
          base_aspect_ratio = .9)
```

```{r n_mito_vs_umis_filter}

n_mito_umi_plot <- ggplot(summary_dat, 
       aes(nUMI, proportion.mito)) + 
  geom_point(aes(color = new_expt_id), 
             size = 0.25) +
  labs(x = "UMIs per cell",
       y = "Proportion of reads\nderived from mitochondria") +
  scale_color_viridis(discrete = TRUE, alpha = 0.33, name = "") +
  guides(colour = guide_legend(ncol = 1, override.aes = list(size = 2))) +
  theme(legend.pos = "top",
        axis.text = element_text(size = 10),
        axis.text.x = element_text(size = 8),
        axis.title = element_text(size = 10),
        legend.text = element_text(size = 6))

save_plot(file.path("fig1", "filtered", "nmito_umi.pdf"), n_mito_umi_plot, 
          base_height = 4, 
          base_aspect_ratio = 1.1)

```

```{r gfp_counts, message = F, eval = T}

n_gfp_plot <- plot_violin(summary_dat, 
            "new_expt_id", "gfp_counts", 
            .fill = "new_expt_id") + 
  labs(y = "GFP expression") +
  theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.pos = "none")

save_plot(file.path("fig1", "filtered", "ngfp.pdf"), n_gfp_plot, 
           base_height = 4, 
          base_aspect_ratio = .9)
```


```{r single_page_fig2}
plts <- list(
  n_cell_plt,
  n_gene_plot,
  n_umi_plt,
  n_mito_plot,
  n_gfp_plot,
  n_mito_umi_plot
)

plt_top <- plot_grid(plotlist = plts[1:6],
          nrow = 3,
          ncol = 3,
         # align = 'hv',
          axis = "tl")


plt <- plot_grid(plt_top, plts[[6]], 
                 nrow = 2, ncol = 1,
                 rel_heights = c(1.5, 1))

save_plot(file.path("fig1", "filtered", "filtered_combined.pdf"), 
          plt_top, base_width = 8, base_height = 11)
```

## Figure 2 code

### By experiments

```{r tsnes}
dir.create("fig2", recursive = TRUE, showWarnings = FALSE)

#tsne_col_pal <- brewer.pal(6, sa"Paired")
#tsne_col_pal <- rev(viridis(6))
 #1,2

sample_tsne <- plot_tsne(atcells, 
                         ident = "new_expt_id",
                         pt.alpha = 0.50,
                         .cols = tsne_col_pal) 

save_plot(file.path("fig2", "tsne_samples_per_expt.pdf"), sample_tsne, 
           base_height = 4, 
           base_aspect_ratio = 1.66)
```

#### all expts
```{r tsne_by_expt}
dir.create(file.path("fig2", "tsne_per_expt"))

expt1 <- rownames(atcells@meta.data)[atcells@meta.data$expt == "expt1"]
expt2 <- rownames(atcells@meta.data)[atcells@meta.data$expt == "expt2"]

plot_expt_tsnes <- function(seurat_obj = atcells, 
                            expt1_cells = expt1,
                            expt2_cells = expt2,
                            expt1_name = "Experiment 1",
                            expt2_name = "Experiment 2",
                            out_name = "all",
                            tsne_cols = tsne_col_pal,
                            outdir = file.path("fig2", "tsne_per_expt")) {
  
    paired_tsne_theme  <- theme(axis.title.x = element_blank(),
            axis.text = element_text(size = 10),
            axis.title = element_text(size = 10),
            legend.pos = "top")
    
    limit_setting <- lims(x = c(min(seurat_obj@dr$tsne@cell.embeddings[, 1]), 
                                max(seurat_obj@dr$tsne@cell.embeddings[, 1])),
                          y = c(min(seurat_obj@dr$tsne@cell.embeddings[, 2]),
                                max(seurat_obj@dr$tsne@cell.embeddings[, 2])))
       
    tsne_expts <- plot_tsne(seurat_obj, 
              ident = "expt",
              pt.alpha = 0.5,
              .cols = tsne_cols, 
              cell_filter = c(expt1_cells, expt2_cells),
              legend_names = c(expt1_name,
                               expt2_name)) +
      paired_tsne_theme
    
    
    tsne_expt1 <- plot_tsne(seurat_obj, 
              ident = "expt",
              pt.alpha = 0.5,
              .cols = tsne_cols[1], 
              cell_filter = expt1_cells,
              legend_names = expt1_name) +
      limit_setting +
      paired_tsne_theme
    
    
    tsne_expt2 <- plot_tsne(seurat_obj, 
              ident = "expt",
              pt.alpha = 0.5,
              .cols = tsne_cols[2], 
              cell_filter = expt2_cells,
              legend_names = expt2_name) +
      limit_setting +
      paired_tsne_theme
    
    plt <- plot_grid(tsne_expt1,
              tsne_expt2,
              tsne_expts, 
              nrow = 1, 
              align = 'hv',
              axis = 'l')
    
    save_plot(file.path(outdir, paste0("tsne_", out_name, "_expts.pdf")), 
              plt,
              ncol = 3, 
              base_height = 3,
              base_aspect_ratio = 1.1)
    
    save_plot(file.path(outdir, paste0("tsne_", out_name, "_expt1.pdf")), 
              tsne_expt1,
              base_height = 3,
              base_aspect_ratio = 1.1)
    
    save_plot(file.path(outdir, paste0("tsne_", out_name, "_expt2.pdf")), 
              tsne_expt2,
              base_height = 3,
              base_aspect_ratio = 1.1)
    
    save_plot(file.path(outdir, paste0("tsne_", out_name, "_expt1+2.pdf")), 
              tsne_expts,
              base_height = 3,
              base_aspect_ratio = 1.1)
    
}

plot_expt_tsnes(atcells)


expt1 <- rownames(atcells@meta.data)[atcells@meta.data$new_expt_id == "ATII expt.1"]
expt2 <- rownames(atcells@meta.data)[atcells@meta.data$new_expt_id == "ATII expt.2" ]

plot_expt_tsnes(atcells,
                expt1_cells = expt1,
                expt2_cells = expt2,
                expt1_name = "naive-ATII\nexpt.1",
                expt2_name = "naive-ATII\nexpt.2",
                out_name = "naive_type_ii",
                tsne_cols = tsne_col_pal[c(1, 2)])
                          
#### naive type i

expt1 <- rownames(atcells@meta.data)[atcells@meta.data$new_expt_id == "Non-ATII epithelial expt.1"]
expt2 <- rownames(atcells@meta.data)[atcells@meta.data$new_expt_id == "Non-ATII epithelial expt.2" ]
plot_expt_tsnes(atcells,
                expt1_cells = expt1,
                expt2_cells = expt2,
                expt1_name = "naive-ATI\nexpt.1",
                expt2_name = "naive-ATI\nexpt.2",
                out_name = "naive_type_i",
                tsne_cols = tsne_col_pal[c(1, 2)])

#### naive type ii injured

expt1 <- rownames(atcells@meta.data)[atcells@meta.data$new_expt_id == "ATII-injured expt.1"]
expt2 <- rownames(atcells@meta.data)[atcells@meta.data$new_expt_id == "ATII-injured expt.2" ]

plot_expt_tsnes(atcells,
                expt1_cells = expt1,
                expt2_cells = expt2,
                expt1_name = "ATII-injured expt.1",
                expt2_name = "ATII-injured expt.2",
                out_name = "injured_type_ii",
                tsne_cols = tsne_col_pal[c(1, 2)])
```

#### all expts split out technical replicates
```{r tsne_by_expt_tech_reps}

atcells <- AddMetaData(atcells, 
                       data.frame(row.names = rownames(atcells@meta.data),
                                  technical_replicate = str_match(atcells@meta.data$sample_names, "([0-9])expt")[, 2]))
                        
atcells <- AddMetaData(atcells,
                       data.frame(row.names = rownames(atcells@meta.data),
                                  experiment_and_tech_rep = str_c(atcells@meta.data$experiment,
                                  ": Technical Replicate ",
                                  atcells@meta.data$technical_replicate)))

dir.create(file.path("fig2", "tsne_per_expt", "technical_replicates"))

expt1 <- rownames(atcells@meta.data)[atcells@meta.data$expt == "expt1"]
expt2 <- rownames(atcells@meta.data)[atcells@meta.data$expt == "expt2"]

plot_expt_tech_tsnes <- function(seurat_obj = atcells, 
                            expt1_cells = expt1,
                            expt2_cells = expt2,
                            expt1_name = paste("Technical replicate", 
                                               c("1", "2")),
                            expt2_name = paste("Technical replicate", 
                                               c("1", "2")),
                            out_name = "all",
                           # tsne_cols = brewer.pal(9, "Paired")[c(6, 5, 2, 1)],
                            tsne_cols = brewer.pal(4, "Set1"),
                            combined_cols = tsne_cols,
                            outdir = file.path("fig2", 
                                               "tsne_per_expt",
                                               "technical_replicates"),
                            save_plots = T) {
  
    paired_tsne_theme  <- theme(axis.title.x = element_blank(),
            axis.text = element_text(size = 10),
            axis.title = element_text(size = 10)) 
    
    limit_setting <- lims(x = c(min(seurat_obj@dr$tsne@cell.embeddings[, 1]), 
                                max(seurat_obj@dr$tsne@cell.embeddings[, 1])),
                          y = c(min(seurat_obj@dr$tsne@cell.embeddings[, 2]),
                                max(seurat_obj@dr$tsne@cell.embeddings[, 2])))
       
    tsne_expts <- plot_tsne(seurat_obj, 
              ident = "experiment_and_tech_rep",
              pt.alpha = 0.5,
              .cols = combined_cols, 
              cell_filter = c(expt1_cells, expt2_cells)) +
      paired_tsne_theme +
      guides(color = guide_legend(nrow = 4 , byrow=TRUE,
                                  label.position = "left",
                                  override.aes = list(size = 5)))
    
    if (save_plots){
      tsne_expts <- tsne_expts  + theme(legend.pos = "none")
    } else{
      tsne_expts <- tsne_expts + theme(legend.pos = "top")
    }
    
    tsne_expt1 <- plot_tsne(seurat_obj, 
              ident = "technical_replicate",
              pt.alpha = 0.5,
              .cols = tsne_cols[1:2], 
              cell_filter = expt1_cells,
              legend_names = expt1_name) +
      limit_setting +
      paired_tsne_theme +
      theme(legend.pos = "none")
    
    
    tsne_expt2 <- plot_tsne(seurat_obj, 
              ident = "technical_replicate",
              pt.alpha = 0.5,
              .cols = tsne_cols[3:4], 
              cell_filter = expt2_cells,
              legend_names = expt2_name) +
      limit_setting +
      paired_tsne_theme +
      theme(legend.pos = "none")
    
    plt <- plot_grid(tsne_expt1,
              tsne_expt2,
              tsne_expts, 
              nrow = 1, 
              align = 'hv',
              axis = 'l')
    
    if(save_plots){
      save_plot(file.path(outdir, paste0("tsne_", out_name, "_expts.pdf")), 
              plt,
              ncol = 3, 
              base_height = 3,
              base_aspect_ratio = 1.1)
    } else{
      return(tsne_expts)
    }
    
}

plt <- plot_expt_tech_tsnes(atcells, save_plots = F)
save_plot(file.path("fig2", 
                    "tsne_per_expt",
                    "technical_replicates", "legend.pdf"), get_legend(plt))

plot_expt_tech_tsnes(atcells, save_plots = T)

expt1 <- rownames(atcells@meta.data)[atcells@meta.data$new_expt_id == "ATII expt.1"]
expt2 <- rownames(atcells@meta.data)[atcells@meta.data$new_expt_id == "ATII expt.2" ]

plot_expt_tech_tsnes(atcells,
                expt1_cells = expt1,
                expt2_cells = expt2,
                out_name = "naive_type_ii",
                combined_cols = brewer.pal(4, "Set1")[c(3, 1)])
                          
#### naive type i

expt1 <- rownames(atcells@meta.data)[atcells@meta.data$new_expt_id == "Non-ATII epithelial expt.1"]
expt2 <- rownames(atcells@meta.data)[atcells@meta.data$new_expt_id == "Non-ATII epithelial expt.2" ]
plot_expt_tech_tsnes(atcells,
                expt1_cells = expt1,
                expt2_cells = expt2,
                expt1_name = "naive-ATI\nexpt.1",
                expt2_name = "naive-ATI\nexpt.2",
                out_name = "naive_type_i")

#### naive type ii injured

expt1 <- rownames(atcells@meta.data)[atcells@meta.data$new_expt_id == "ATII-injured expt.1"]
expt2 <- rownames(atcells@meta.data)[atcells@meta.data$new_expt_id == "ATII-injured expt.2" ]

plot_expt_tech_tsnes(atcells,
                expt1_cells = expt1,
                expt2_cells = expt2,
                expt1_name = "ATII-injured expt.1",
                expt2_name = "ATII-injured expt.2",
                out_name = "injured_type_ii")


## readme
write_lines(c("tsnes are colored to show each technical replicate", 
              "the first column is the first experiment",
              "the second column is the second experiment",
              "the third column is combined",
              "the legend.pdf contains the legend"),
            file.path("fig2", "tsne_per_expt","technical_replicates", "README.txt"))

```


### By sample type

Group each experiment together and plot naive non-atii epithelial, naive at II, injured at II, and all

```{r by_sample_grouped, fig.width = 12}


mdata <- atcells@meta.data

naive_non_atii_epi <- rownames(mdata)[mdata$sample_type == "Naive Non-AT-II Epithelial"]
naive_atii <- rownames(mdata)[mdata$sample_type == "Naive AT-II"]
injured_atii <- rownames(mdata)[mdata$sample_type =="Injured AT-II"]

cols <- brewer.pal(3, "Set1")
plt <- plot_tsne(atcells, ident = "sample_type",
               .cols = rev(cols)[1:2], pt.alpha = 0.50, 
               cell_filter = c(naive_non_atii_epi,
          naive_atii)) + 
  theme(legend.pos = "top")

save_plot(file.path("fig2","tsne_naive_only.pdf"), plt, 
          base_height = 6, base_aspect_ratio = 1.33)

plts <- map2(list(naive_non_atii_epi,
          naive_atii,
          injured_atii),
     rev(cols),
     ~plot_tsne(atcells, ident = "sample_type", cell_filter = .x, 
               .cols = .y, pt.alpha = 0.50) + theme(legend.pos = "top"))

plts[[length(plts) + 1]] <- plot_tsne(atcells, ident = "sample_type", 
               .cols = cols, pt.alpha = 0.50) + theme(legend.pos = "top")

plt <- plot_grid(plotlist = plts, nrow = 1)
save_plot(file.path("fig2","tsne_per_sampletype.pdf"), plt, 
                    nrow = 1, base_height = 5.5, base_aspect_ratio = 4)

sample_type_cells <- lst(naive_non_atii_epi,
          naive_atii,
          injured_atii)

plts <- pmap(list(sample_type_cells,
     rev(cols),
     list(c("Experiment 1",
       "Experiment 2",
       "Combined"), rep("", 3), rep("", 3))), 
     function(.x, .y, .z){
     expt1 <- rownames(atcells@meta.data)[atcells@meta.data$experiment == "Experiment 1"]
     expt2 <- rownames(atcells@meta.data)[atcells@meta.data$experiment == "Experiment 2"]
     
     expt1 <- intersect(.x, expt1)
     expt2 <- intersect(.x, expt2)
     
     plt_theme <- theme(legend.pos = "none",
                        axis.line = element_blank(),
                        axis.title = element_blank(),
                        axis.text = element_blank(),
                        axis.ticks = element_blank(),
                        plot.background = element_rect(colour = "grey", 
                                                       linetype = "solid", size = 0.5))
     
     plt1 <- plot_tsne(atcells, ident = "sample_type", cell_filter = expt1, 
               .cols = .y, pt.alpha = 0.50) + labs(title = .z[1]) + plt_theme
     plt2 <- plot_tsne(atcells, ident = "sample_type", cell_filter = expt2, 
               .cols = .y, pt.alpha = 0.50)  + labs(title = .z[2]) + plt_theme
     plt3 <- plot_tsne(atcells, ident = "sample_type", cell_filter = .x, 
               .cols = .y, pt.alpha = 0.50)  + labs(title = .z[3]) + plt_theme
     plt <- plot_grid(plt1, plt2, plt3, nrow = 1)
     plt
})

plt <- plot_grid(plotlist = plts, ncol = 1,  labels = c("Naive Non-AT-II Epithelial",
                                                        "Naive AT-II",
                                                        "Injured AT-II"),
                 label_size = 8, label_x = 0, label_y = 1)
save_plot(file.path("fig2","tsne_per_sampletype_expt.pdf"), plt, 
          nrow = 1, base_height = 7, base_aspect_ratio = 1)


###### No labels ######
plts <- pmap(list(sample_type_cells,
     rev(cols)), 
     function(.x, .y){
     expt1 <- rownames(atcells@meta.data)[atcells@meta.data$experiment == "Experiment 1"]
     expt2 <- rownames(atcells@meta.data)[atcells@meta.data$experiment == "Experiment 2"]
     
     expt1 <- intersect(.x, expt1)
     expt2 <- intersect(.x, expt2)
     
     plt_theme <- theme(legend.pos = "none",
                        axis.line = element_blank(),
                        axis.title = element_blank(),
                        axis.text = element_blank(),
                        axis.ticks = element_blank(),
                        plot.background = element_rect(colour = "grey", 
                                                       linetype = "solid", size = 0.5))
     
     plt1 <- plot_tsne(atcells, ident = "sample_type", cell_filter = expt1, 
               .cols = .y, pt.alpha = 0.50) + plt_theme
     plt2 <- plot_tsne(atcells, ident = "sample_type", cell_filter = expt2, 
               .cols = .y, pt.alpha = 0.50)   + plt_theme
     plt3 <- plot_tsne(atcells, ident = "sample_type", cell_filter = .x, 
               .cols = .y, pt.alpha = 0.50)  + plt_theme
     
     plt <- plot_grid(plt1, plt2, plt3, nrow = 1)
     plt
})

plt <- plot_grid(plotlist = plts, ncol = 1)
save_plot(file.path("fig2","tsne_per_sampletype_expt_nolabels.pdf"), plt, 
          nrow = 1, base_height = 7, base_aspect_ratio = 1)
```

### Markers

#### tnse by sample type function

```{r, tsne_by_sample_type}
library(gridExtra)
g_legend <- function(a.gplot){
    tmp <- ggplot_gtable(ggplot_build(a.gplot))
    leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
    legend <- tmp$grobs[[leg]]
    legend
}


tsne_by_sample_type <- function(seurat_obj, .gene, cell_sets, 
                               top_label = NULL, 
                               limit_gene_exp = NULL, 
                               resize_tsne = FALSE,
                               ...){
      
    if (is.null(limit_gene_exp )){
      limit_gene_exp <- c(0, max(FetchData(seurat_obj, 
                                         .gene, 
                                         cells.use = unique(unlist(cell_sets)))))
    }
    #if(limit_gene_exp[2] == 0) return()
    
    if(resize_tsne){
      all_cells <- unique(unlist(cell_sets)) 
      sub_obj <- SubsetData(seurat_obj, cells.use = all_cells)
    } else{
      sub_obj <- seurat_obj
    }

    limit_setting <- lims(x = c(min(sub_obj@dr$tsne@cell.embeddings[, 1]), 
                                max(sub_obj@dr$tsne@cell.embeddings[, 1])),
                          y = c(min(sub_obj@dr$tsne@cell.embeddings[, 2]),
                                max(sub_obj@dr$tsne@cell.embeddings[, 2])))
    
     plt_theme <- theme(axis.line = element_blank(),
                        axis.title = element_blank(),
                        axis.text = element_blank(),
                        axis.ticks = element_blank(),
                        plot.background = element_rect(colour = "grey", 
                                                       linetype = "solid", size = 0.5))
     
     if(is.null(top_label)){ top_label <- rep("", length(cell_sets))}
     
     plts = list()
     for (i in seq_along(1:length(cell_sets))){
         plts[[i]] <- plot_feature(seurat_obj, gene = .gene, 
                                   cell_filter = cell_sets[[i]],
                                   pt.alpha = 1, max_y = limit_gene_exp, ...) + 
           limit_setting +
           labs(title = top_label[i]) +
           theme(legend.position = "none") +
           plt_theme
         
       if(i == length(cell_sets)){
         tmp <- plot_feature(seurat_obj, gene = .gene, 
                                   cell_filter = cell_sets[[i]],
                                   pt.alpha = 1, max_y = limit_gene_exp, ...) + 
           limit_setting +
           labs(title = top_label[i]) +
           plt_theme
         plts[[i + 1]] <- g_legend(tmp)
       }
     }
     
     plot_grid(plotlist = plts, nrow = 1,
               rel_widths = c(rep(1, length(cell_sets)), .33))
}


```
#### ciliary

```{r}
dir.create(file.path("fig2", "tsne_cell_markers_all_cells"))
```

```{r ciliary}
ciliary_markers <- c("Foxj1", 
             "Ccdc39", 
             "Tubb4a",
             "Ccdc113")

plts <- map(ciliary_markers, 
            ~plot_feature(atcells, 
                          gene = .x,
                          pt.alpha = 0.75))

plt <- plot_grid(plotlist = plts, ncol = 2)
save_plot(file.path("fig2", "tsne_cell_markers_all_cells", "tsne_ciliary.pdf"),
          plt,
          ncol = 2, 
          nrow = 2,
          base_height = 4, 
          base_aspect_ratio = 1.33)
```
#### club

```{r club}
club_markers <- c("Scgb1a1", "Scgb3a1", "Scgb3a2", "Chad")

plts <- map(club_markers, 
            ~plot_feature(atcells, 
                          gene = .x, pt.alpha = 1))

plt <- plot_grid(plotlist = plts, ncol = 2)
save_plot(file.path("fig2", "tsne_cell_markers_all_cells", "tsne_club.pdf"),
          plt,
          ncol = 2, 
          nrow = 2, 
          base_height = 4, 
          base_aspect_ratio = 1.33)


```
#### basal
```{r basal}
basal_markers <- c("Krt5", "Trp63", "Krt14", "Krt15")

plts <- map(basal_markers, 
            ~plot_feature(atcells, gene = .x, pt.alpha = 1))

plt <- plot_grid(plotlist = plts, ncol = 2)
save_plot(file.path("fig2", "tsne_cell_markers_all_cells", "tsne_basal.pdf"),
          plt,
          ncol = 2, 
          nrow = 2, 
          base_height = 4, 
          base_aspect_ratio = 1.33)
```

#### At1
```{r at1}
at1_markers <- c("Pdpn", "Emp2", "Hopx", "Akap5", "Aqp5", "Clic5")
plts <- map(at1_markers, 
            ~plot_feature(atcells, gene = .x, pt.alpha = 1))

plt <- plot_grid(plotlist = plts, ncol = 2)
save_plot(file.path("fig2", "tsne_cell_markers_all_cells", "tsne_at1.pdf"),
          plt,
          ncol = 2, 
          nrow = 2, 
          base_height = 4, 
          base_aspect_ratio = 1.33)
```

#### At2
```{r at2}
at2_markers <- c("Sftpc", "Lamp3", "Lyz2", "Abca3", "Sftpa1")

plts <- map(at2_markers, 
            ~plot_feature(atcells, gene = .x, pt.alpha = 1))

plt <- plot_grid(plotlist = plts, ncol = 2)
save_plot(file.path("fig2", "tsne_cell_markers_all_cells", "tsne_at2.pdf"),
          plt,
          ncol = 2, 
          nrow = 2, 
          base_height = 4, 
          base_aspect_ratio = 1.33)
```

#### Macrophages
```{r macs}
#ptprc = cd45
#
macs_markers <- c("Ptprc", "Cd68", "Siglecf")

plts <- map(macs_markers, 
            ~plot_feature(atcells, gene = .x, pt.alpha = 1))

plt <- plot_grid(plotlist = plts, ncol = 2)
save_plot(file.path("fig2", "tsne_cell_markers_all_cells", "tsne_macs.pdf"),
          plt,
          ncol = 2, 
          nrow = 2, 
          base_height = 4, 
          base_aspect_ratio = 1.33)
```

#### Eosinophils

```{r eos}
#ptprc = cd45
#
eos_markers <- c("Ptprc", "Fcer1g")

plts <- map(eos_markers, 
            ~plot_feature(atcells, gene = .x, pt.alpha = 1))

plt <- plot_grid(plotlist = plts, ncol = 2)
save_plot(file.path("fig2", "tsne_cell_markers_all_cells", "tsne_eosionphils.pdf"),
          plt,
          ncol = 2, 
          nrow = 1, 
          base_height = 4, 
          base_aspect_ratio = 1.33)
```

#### fibroblasts
```{r fibs}
# fsp1 = s100a4
fibs_markers <- c("Col1a1", 
                  "Col3a1",
                  "Col1a2",
                  "Vim",
                  "S100a4")

plts <- map(fibs_markers, 
            ~plot_feature(atcells, gene = .x, pt.alpha = 1))

plt <- plot_grid(plotlist = plts, ncol = 2)
save_plot(file.path("fig2", "tsne_cell_markers_all_cells", "tsne_fibroblasts.pdf"),
          plt,
          ncol = 2, 
          nrow = 3, 
          base_height = 4, 
          base_aspect_ratio = 1.33)
```

#### endothelial

```{r endos}
# CD141 = Thbd
# VCAM = Vcam1
endos_markers <- c("Thbd", 
                  "Vcam1")

plts <- map(endos_markers, 
            ~plot_feature(atcells, gene = .x, pt.alpha = 1))

plt <- plot_grid(plotlist = plts, ncol = 2)
save_plot(file.path("fig2", "tsne_cell_markers_all_cells", "tsne_endothelial.pdf"),
          plt,
          ncol = 2, 
          nrow = 1, 
          base_height = 4, 
          base_aspect_ratio = 1.33)
```

#### dendritic

```{r dend}
# CD11c = Itgax
# CD45 = Ptprc
dend_markers <- c("Itgax", 
                  "Ptprc")

plts <- map(dend_markers, 
            ~plot_feature(atcells, gene = .x, pt.alpha = 1))

plt <- plot_grid(plotlist = plts, ncol = 2)
save_plot(file.path("fig2", "tsne_cell_markers_all_cells", "tsne_dendritic.pdf"),
          plt,
          ncol = 2, 
          nrow = 1, 
          base_height = 4, 
          base_aspect_ratio = 1.33)
```

#### ly6

```{r ly6}
#CD3 = Cd3e
#CD4 = Cd4
#CD8 = Cd8a
#CD19 = Cd19
#B220 = Ptprc
#Ly6G = Ly6g

# CD141 = Thbd
# VCAM = Vcam1
ly6_markers <- c("Cd3e",
                 "Cd4",
                 "Cd8b1",
                 "Cd19",
                 "Ptprc",
                 "Ly6g")

plts <- map(ly6_markers, 
            ~plot_feature(atcells, gene = .x, pt.alpha = 1))

plt <- plot_grid(plotlist = plts, ncol = 2)
save_plot(file.path("fig2", "tsne_cell_markers_all_cells", "tsne_ly6g.pdf"),
          plt,
          ncol = 2, 
          nrow = 3, 
          base_height = 4, 
          base_aspect_ratio = 1.33)
```

### by naive type 

#### tsne
```{r all_cells_by_sample_type}
dir.create(file.path("fig2", "tsne_naive_by_sample_type"))
dir.create(file.path("fig2", "tsne_naive_by_sample_type", "tsne_plots"))
dir.create(file.path("fig2", "tsne_naive_by_sample_type", "tsne_plots", "no_labels"))

write_lines(
  "tsne_naive_by_sample_type\tTSNE and violin plots split by the naive sample type for each cell\tCd19 was removed as it had no expression",
  file.path("fig2", "README.txt"), append = F)

cell_markers <- lst(
  ciliary_markers,
  club_markers,
  endos_markers, 
  fibs_markers, 
  eos_markers, 
  macs_markers, 
  at2_markers, 
  at1_markers, 
  basal_markers,
  dend_markers,
  ly6_markers
)

naive_cell_types <- list("naive_non_atii_epi" = sample_type_cells[[1]],
                         "naive_atii" = sample_type_cells[[2]],
                         "all_naive" = c(sample_type_cells[[1]],
                                         sample_type_cells[[2]]))

### with labels
tsne_plts <- map2(cell_markers,
     names(cell_markers),
    function(x, y){
  plts <- list()
  plts[[1]] <- tsne_by_sample_type(atcells, 
                    .gene = x[1],
                    cell_sets = naive_cell_types,
                    top_label = names(naive_cell_types))
  
  plts_no_label <- map(x[2:length(x)], 
           ~tsne_by_sample_type(atcells, 
                    .gene = .x,
                    cell_sets = naive_cell_types))
  plt <- plot_grid(plotlist = c(plts, plts_no_label), ncol = 1)

  save_plot(file.path("fig2", 
                      "tsne_naive_by_sample_type", 
                      "tsne_plots", 
                      paste0("tsne_", y, ".pdf")),
          plt,
          ncol = 1,
          nrow = length(x),
          base_height = 2.75, 
          base_aspect_ratio = 3)
  plt
})

### without labels
map2(cell_markers,
     names(cell_markers),
    function(x, y){
  plts <- list()
  plts[[1]] <- tsne_by_sample_type(atcells, 
                    .gene = x[1],
                    cell_sets = naive_cell_types)

  plts_no_label <- map(x[2:length(x)], 
           ~tsne_by_sample_type(atcells, 
                    .gene = .x,
                    cell_sets = naive_cell_types))
       
  plt <- plot_grid(plotlist = c(plts, plts_no_label), ncol = 1)

  save_plot(file.path("fig2", 
                      "tsne_naive_by_sample_type", 
                      "tsne_plots", 
                      "no_labels", 
                      paste0("tsne_", y, ".pdf")),
          plt,
          ncol = 1,
          nrow = length(x),
          base_height = 2.75, 
          base_aspect_ratio = 3)
})

```

#### violin plots

```{r}
# merge populations from naive type ii
ids_to_merge <- c("Naive Type II",
                  "Cell Cycle Arrest Type II",
                  "Injured Type II",
                  "Transdifferentiating Type II",
                  "Proliferating Type II")

mdata <- atcells@meta.data[, 
                           c("labeled_clusters", "sample_type")] %>% 
  tibble::rownames_to_column("cell") 

mdata <- filter(mdata, cell %in% unique(unlist(naive_cell_types)),
                !(sample_type == "Naive Non-AT-II Epithelial" &
                    labeled_clusters %in% ids_to_merge))

mdata <- mutate(mdata,
                new_id = ifelse(sample_type == "Naive AT-II" & 
                                  labeled_clusters %in% ids_to_merge,
                                "Naive AT-II",
                                 ifelse(labeled_clusters == "Naive Type I",
                                              "Naive AT-I",
                                              labeled_clusters)))

dir.create(file.path("fig2", 
                     "tsne_naive_by_sample_type", 
                     "violin_plots"))
imap(cell_markers,
    function(x, y){
     plts <- map(x,
       function(x){
       dat <- FetchData(atcells, x) %>% 
         as.data.frame() %>%
         tibble::rownames_to_column("cell")
       
       dat <- inner_join(dat, mdata, by = "cell")
       max_expr <- max(dat[[x]])
       p <- plot_violin(dat, "new_id", x, .scale = "count", 
                   single_col = brewer.pal(11, "RdGy")[7]) 
       if (max_expr != 0){
         p <- p + geom_jitter(size = 0.1, alpha = 0.2)
       }
       p + theme(axis.title.x = element_blank())
     })
    plt <- plot_grid(plotlist = plts, ncol = 1)
    save_plot(file.path("fig2", 
                        "tsne_naive_by_sample_type", "violin_plots", 
                        paste0(y, ".pdf")), 
              ncol = 1, 
              nrow = length(x),
              plt, 
              base_height = 2.5, 
              base_aspect_ratio = 1.1)
 })



```

```{r per_cell_type}
non_type2 <- rownames(atcells@meta.data)[str_detect(atcells@meta.data$new_expt_id, 
                                                    "^Non-ATII epithelial")]
type2 <- rownames(atcells@meta.data)[str_detect(atcells@meta.data$new_expt_id, 
                                                "^ATII expt")]

paired_tsne_theme  <- theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10))

plot_grouped_features <- function(seurat_obj, 
                                  gene_to_plot = "Krt5",
                                  grp_1_cells = non_type2,
                                  grp_2_cells = type2,
                                  grp_1_title = "Naive Non-ATII",
                                  grp_2_title = "Naive ATII",
                                  all_title = "All Cells",
                                  plt_theme = paired_tsne_theme,
                                  out_dir = "fig2") {
  
    limit_gene_exp <- c(0, max(FetchData(seurat_obj, 
                                         gene_to_plot, 
                                         cells.use = c(grp_1_cells,
                                                       grp_2_cells))))
  
    limit_setting <- lims(x = c(min(seurat_obj@dr$tsne@cell.embeddings[, 1]), 
                                max(seurat_obj@dr$tsne@cell.embeddings[, 1])),
                          y = c(min(seurat_obj@dr$tsne@cell.embeddings[, 2]),
                                max(seurat_obj@dr$tsne@cell.embeddings[, 2])))
    
    tsne_all <- plot_feature(seurat_obj, 
              gene = gene_to_plot,
              cell_filter = NULL,
              max_y = limit_gene_exp) + 
      limit_setting + 
      labs(title = all_title) +
      plt_theme
    
    tsne_grp1 <- plot_feature(seurat_obj, 
              gene = gene_to_plot,
              cell_filter = grp_1_cells,
              max_y = limit_gene_exp) + 
      limit_setting + 
      labs(title = grp_1_title) +
      plt_theme
    
    tsne_grp2 <- plot_feature(seurat_obj, 
              gene = gene_to_plot,
              cell_filter = grp_2_cells,
              max_y = limit_gene_exp) + 
      limit_setting + 
      labs(title = grp_2_title) +
      plt_theme
    
    plt <- plot_grid(tsne_grp1,
              tsne_grp2, 
              tsne_all,
              nrow = 1, 
              align = 'hv',
              axis = 'l')
    
    grp_1_title <- str_replace_all(grp_1_title, "( |-)", "_")
    grp_2_title <- str_replace_all(grp_2_title, "( |-)", "_")
    
    save_plot(file.path(out_dir, paste0(gene_to_plot, "_tsne_all_plots.pdf")), 
              plt,
              ncol = 3, 
              base_height = 3,
              base_aspect_ratio = 1.25)
    
    save_plot(file.path(out_dir, paste0(gene_to_plot, "_tsne_all.pdf")), 
              tsne_all,
              base_height = 3,
              base_aspect_ratio = 1.25)
    
    save_plot(file.path(out_dir, paste0(gene_to_plot, "_tsne_", grp_1_title, ".pdf")), 
              tsne_grp1,
              base_height = 3,
              base_aspect_ratio = 1.25)
    
    save_plot(file.path(out_dir, paste0(gene_to_plot, "_tsne_", grp_2_title, ".pdf")), 
              tsne_grp2,
              base_height = 3,
              base_aspect_ratio = 1.25)
}

cell_markers <- c(
  ciliary_markers,
  club_markers,
  endos_markers, 
  fibs_markers, 
  eos_markers, 
  macs_markers, 
  at2_markers, 
  at1_markers, 
  basal_markers,
  dend_markers,
  ly6_markers
)

dir.create(file.path("fig2", "misc", "tsne_markers_by_cell_type"), 
           recursive = T,
           showWarnings = FALSE)

map(cell_markers, function(x){
  out_path <- file.path("fig2", 
                        "misc",
                       "tsne_markers_by_cell_type",
                       x)
  dir.create(out_path,
             recursive = T,
             showWarnings = FALSE)
  
  plot_grouped_features(atcells,
                      gene_to_plot = x,
                      out_dir = out_path)})

```

### cell markers matched tsne + violin

```{r tsne_violins}

plot_violin_tsne <- function(seurat_obj, 
                             .ident = "labeled_clusters",
                             .gene = "Hopx",
                             cell_filter = NULL){
  
  gene_dat <- FetchData(seurat_obj, .gene) %>% 
      as.data.frame() %>% 
      tibble::rownames_to_column("cell")
  
  if (!is.null(cell_filter)){
    gene_dat <- dplyr::filter(gene_dat,
                                cell %in% cell_filter)
  }
  
  mdata <- seurat_obj@meta.data %>%
    tibble::rownames_to_column("cell")
  
  plot_dat <- left_join(gene_dat, mdata, by = "cell")
  vln_plt <- plot_violin(plot_dat, 
              .x = .ident,
              .y = .gene,
              .fill = .ident,
              .scale = "width",
              jitter = T,
              .alpha = 0.5,
              single_col = brewer.pal(11, "RdGy")[7]) +
    theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.pos = "none")
  
  tsne_plt <- plot_feature(seurat_obj, 
              gene = .gene,
              cell_filter = cell_filter,
              pt.alpha = 1) +
    guides(color = guide_colorbar(barwidth = 0.5, barheight = 5)) +
    theme(axis.text = element_text(size = 10),
          axis.title = element_text(size = 10),
          legend.title = element_text(size = 10),
          legend.text = element_text(size = 10))
  
  empty_plot <- ggplot(tsne_plt$data) +
    geom_blank()
  
  plt <- plot_grid(tsne_plt,
            empty_plot,
            ncol = 1,
            rel_heights = c(4, 1))
  
  plt <- plot_grid(plt,
                   vln_plt,
                   nrow = 1, 
                   axis = 'h',
                   align = 'bl',
                   rel_widths = c(1.5, 1),
                   rel_heights = c(1.5, 1))
  plt
}


dir.create(file.path("fig2", "all_cells_violin_tsnes"))

map(cell_markers, function(x){
  out_name <- file.path("fig2", 
                       "all_cells_violin_tsnes",
                       paste0(x, ".pdf"))
  
  plt <- plot_violin_tsne(atcells, .gene = x)
  save_plot(out_name, plt,
          base_height = 4, 
          base_aspect_ratio = 1.5)})
```

#### violin per cell type

```{r vln_cell_type}

plot_violin_gene <- function(seurat_obj, 
                             .ident = "labeled_clusters",
                              .genes = "Hopx"){
  
  gene_dat <- FetchData(seurat_obj, .genes) %>% 
      as.data.frame() %>% 
      tibble::rownames_to_column("cell")
  
  mdata <- seurat_obj@meta.data %>% tibble::rownames_to_column("cell")
  
  plot_dat <- left_join(mdata, gene_dat, by = "cell")
  
  plts <- map(.genes, 
              ~plot_violin(plot_dat, 
              .x = .ident,
              .y = .x,
              .fill = .ident,
              .scale = "width",
              jitter = T,
              .alpha = 0.5,
              single_col = brewer.pal(11, "RdGy")[7]) +
    theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 10),
        legend.pos = "none")
  )
  plot_grid(plotlist = plts)
}

cell_markers_lst <- lst(
  ciliary_markers,
  club_markers,
  endos_markers, 
  fibs_markers, 
  eos_markers, 
  macs_markers, 
  at2_markers, 
  at1_markers, 
  basal_markers,
  dend_markers,
  ly6_markers
)

dir.create(file.path("fig2", "all_clusters_violin_plots"))
imap(cell_markers_lst, 
     function(x, y) {
       plts <- plot_violin_gene(atcells, .genes = x)
       save_plot(file.path("fig2", 
                           "all_clusters_violin_plots", 
                           paste0(y, ".pdf")), 
                 plts,
                 base_height = 6, 
                 base_aspect_ratio = 1.25)})
```

## Figure 3 code

exclude contaminating populations (Club, ciliated, basal, endothelial,
macrophage, fibroblasts)

TSNE plots and graphs of key proliferation, cell cycle arrest, ATII, ATI cell markers shown w/ all
cells except excluded populations on TSNE and also on injured vs. uninjured separately 
identify populations
a. Proliferation: mKi67, PCNA, CDK1, Topo2, Cyclin B1, Cyclin A2, Rrm1, E2F8, E2F7, E2F2,
E2F1
b. Cell Cycle Arrest: p15, p53, Cdk4, Cyclin D1, Cyclin D1 BP, E2F4, E2F5
c. ATII-derived ATI: AQP5, Pdpn, EMP2, Ager, hopx, Cav1, other top ATI cell markers
(IGFBP2, Rtkn2, Pxdc1), ATII cell markers: Lamp3, C5a, LPCAT, SPA/B/C/D.


### TD markers

```{r td_markers}
dir.create("fig4")
dir.create(file.path("fig3", "all_cells_violin_tsnes"), 
           recursive = T, showWarnings = F)

cell_types_to_keep <- c(
  "Injured Type II",
  "Proliferating Type II",
  "Transdifferentiating Type II",
  "Cell Cycle Arrest Type II",
  "Naive Type I",
  "Naive Type II"
)

idx <- atcells@meta.data$labeled_clusters %in% cell_types_to_keep
cells_to_keep <- rownames(atcells@meta.data)[idx]
cells_to_keep_data <- atcells@meta.data[idx, ]

injured_cells <- rownames(cells_to_keep_data[!str_detect(cells_to_keep_data$labeled_clusters, "Naive"), ])
naive_cells <- rownames(cells_to_keep_data[str_detect(cells_to_keep_data$labeled_clusters,
                                               "Naive"), ])

```


Couldn't figure out the mouse gene symbol for Cyclin D1 BP 
```{r plt_fxn}
readme_file <- file.path("fig3", "all_cells_violin_tsnes", "README")
if(file.exists(readme_file)){
  unlink(readme_file)
}
write_lines("A is uninjured and injured together", readme_file)
write_lines("B is uninjured ", readme_file)
write_lines("C is injured ", readme_file)

cells_to_plot <- list(cells_to_keep,
                   naive_cells,
                   injured_cells)
  
new_cell_markers <- list(
  proliferation = c( 
    "Mki67", 
    "Pcna", 
    "Cdk1", 
    "Top2a", 
    "Ccnb1", 
    "Ccna2", 
    "Rrm1", 
    "E2f8", 
    "E2f7", 
    "E2f2",
    "E2f1"),
  cell_cycle_arrest = c(
     "Cdkn2b", 
     "Trp53", 
     "Cdk4", 
     "Ccnd1", 
     "E2f4", 
     "E2f5"
  ),
  transdiff = c(
    "Aqp5",
    "Pdpn",
    "Emp2",
    "Ager",
    "Hopx",
    "Cav1",
    "Igfbp2",
    "Rtkn2",
    "Pxdc1",
    "Lamp3",
    "C5ar1",
    "Lpcat1",
    "Sftpa1",
    "Sftpb",
    "Sftpc",
    "Sftpd"
  )
)

violin_tsne_cell_split <- function(gene, outdir,
                                   cells_to_plot = NULL){
  output <- file.path("fig3", "all_cells_violin_tsnes", 
                      outdir, str_c(gene, ".pdf"))

  plts <- map(cells_to_plot, 
            ~plot_violin_tsne(atcells, 
                              .gene = gene,
                               cell_filter = .x))

  plt <- plot_grid(plotlist = plts, 
                 nrow = 3,
                 rel_heights = 4,
                 rel_widths = 6,
                 labels = "AUTO")

  save_plot(output, plt,
          base_height = 4, 
          base_aspect_ratio = 1.5,
          nrow = 3)
}

map2(new_cell_markers,
     names(new_cell_markers),
  function(genes, outdir){
  dir.create(file.path("fig3", "all_cells_violin_tsnes", outdir), showWarnings = F)
  map(genes, ~violin_tsne_cell_split(.x, outdir, cells_to_plot = cells_to_plot))
})
```



### tgfb

```{r tgfb, eval = T}

tgf_markers <- list(
   tgfbeta = c(
     "Tgfbr1", 
     "Tgfbr2", 
     "Tgfb1", 
     "Tgfb2", 
     "Tgfb3", 
     "Ltbp1", 
     "Ltbp2", 
     "Tgfbi", 
     "Smad7", 
     "Smurf1", 
     "Itgb6", 
     "Itgav", 
     "Itga5",
     "Tgif1", 
     "Smad2", 
     "Ctgf", 
     "Acta2", 
     "Col1a1", 
     "Fn1", 
     "Serpine1",
     "Nog",
     "Gadd45b",
     "Fst",
     "Pdgfb",
     "Ptk2b",
     "Nfib",
     "Nfkbia",
     "Smad3", 
     "Smad4", 
     "Cdh1", 
     "Vim"),
   other = c(
     "Igfbp2",
     "Sox9",
     "Id2",
     "Foxp2"
   )
)

injured_cell_types_to_keep <- c(
  "Injured Type II",
  "Proliferating Type II",
  "Transdifferentiating Type II",
  "Cell Cycle Arrest Type II"
)

mdata <- atcells@meta.data

filter_idx <- (mdata$sample_type == "Naive Non-AT-II Epithelial") &
  (mdata$labeled_clusters %in% "Naive Type I")
naive_non_atii_epi <- rownames(mdata)[filter_idx]
naive_atii <- rownames(mdata)[mdata$sample_type == "Naive AT-II" &
                                  mdata$labeled_clusters %in% cell_types_to_keep]
injured_atii <- rownames(mdata)[mdata$sample_type =="Injured AT-II" &
                                  mdata$labeled_clusters %in% injured_cell_types_to_keep]

naive_injured_cell_types <- list(
  "Injured AT-II Derived" = injured_atii,
  "Naive AT-II" = naive_atii,
  "Naive AT-I" = naive_non_atii_epi,
  "Combined" = c(naive_non_atii_epi, naive_atii, injured_atii))

  dir.create(file.path("fig4", "all_cells_tsne_plots", "tgfbeta"), 
             showWarnings = F, recursive = T)
  plts <- map(tgf_markers$tgfbeta, 
      ~plot_feature(atcells, 
                    gene = .x,
                    cell_filter = naive_injured_cell_types$Combined,
                    pt.alpha = 1) + 
        theme_cowplot(font_size = 22))
          
  map2(plts, 
      tgf_markers$tgfbeta,
      ~save_plot(filename = file.path("fig4", "all_cells_tsne_plots",
                                      "tgfbeta",
                                     paste0(.y, ".pdf")),
                .x,
                base_aspect_ratio = 1.33))


dir.create(file.path("fig4", "all_injured_cells_tsne_plots",
                     "tgfbeta"), 
             showWarnings = F, recursive = T)
  plts <- map(tgf_markers$tgfbeta, 
      ~plot_feature(atcells, 
                    gene = .x,
                    cell_filter = naive_injured_cell_types$`Injured AT-II Derived`,
                    pt.alpha = 1)  + 
        theme_cowplot(font_size = 22))
  map2(plts, 
      tgf_markers$tgfbeta,
      ~save_plot(filename = file.path("fig4", "all_injured_cells_tsne_plots",
                                      "tgfbeta",
                                     paste0(.y, ".pdf")),
                .x,
                base_aspect_ratio = 1.33))

  dir.create(file.path("fig4", "all_cells_tsne_plots",
                     "other"), 
             showWarnings = F, recursive = T)
  plts <- map(tgf_markers$other, 
      ~plot_feature(atcells, 
                    gene = .x,
                    cell_filter = naive_injured_cell_types$Combined,
                    pt.alpha = 1)  + 
        theme_cowplot(font_size = 22))
  map2(plts, 
      tgf_markers$other,
      ~save_plot(filename = file.path("fig4", "all_cells_tsne_plots",
                                      "other",
                                     paste0(.y, ".pdf")),
                .x,
                base_aspect_ratio = 1.33))
  
  
  dir.create(file.path("fig4", "all_injured_cells_tsne_plots",
                     "tgfbeta", "no_legend_label"), 
             showWarnings = F, recursive = T)
  plts <- map(tgf_markers$tgfbeta, 
      ~plot_feature(atcells, 
                    gene = .x,
                    cell_filter = naive_injured_cell_types$`Injured AT-II Derived`,
                    pt.alpha = 1, legend_title = "")  + 
        theme_cowplot(font_size = 22))
  map2(plts, 
      tgf_markers$tgfbeta,
      ~save_plot(filename = file.path("fig4", "all_injured_cells_tsne_plots",
                                      "tgfbeta", "no_legend_label",
                                     paste0(.y, ".pdf")),
                .x,
                base_aspect_ratio = 1.33))
  
    dir.create(file.path("fig4", "all_injured_cells_tsne_plots",
                     "tgfbeta"), 
             showWarnings = F, recursive = T)
  plts <- map(tgf_markers$tgfbeta, 
      ~plot_feature(atcells, 
                    gene = .x,
                    cell_filter = naive_injured_cell_types$`Injured AT-II Derived`,
                    pt.alpha = 1)  + 
        theme_cowplot(font_size = 22))
  map2(plts, 
      tgf_markers$tgfbeta,
      ~save_plot(filename = file.path("fig4", "all_injured_cells_tsne_plots",
                                      "tgfbeta",
                                     paste0(.y, ".pdf")),
                .x,
                base_aspect_ratio = 1.33))
```

### Notch wnt signaling

11.	TSNE plots and/or heatmaps and/or graphs of b-catenin, HIF signaling, Notch Signaling
a.	Notch: Jag1, Sox9, Edn1, Efnb1, Efna1, Nes, Rnd1, Flt1, Snw1, Id2, Jun, Cx3cl1, Hes1, Sgpl1, Kalrn, Nampt, Crebbp, Runx1, Tec, Notch1,2,3,4, DLL1, Rbpj, Mib1, c-myb, all genes together?  (cloupe allows you to show TSNEs of expression of 10 genes simulatenously, I guess it is average expression level)

```{r wnt_notch, eval = T}

notch_markers <- list(
   notch = c(
    "Jag1",
    "Sox9", 
  	"Edn1", 
  	"Efnb1", 
  	"Efna1", 
  	"Nes", 
  	"Rnd1", 
  	"Flt1", 
  	"Snw1", 
  	"Id2", 
  	"Jun", 
  	"Cx3cl1", 
  	"Hes1", 
  	"Sgpl1", 
  	"Kalrn", 
  	"Nampt", 
  	"Crebbp", 
  	"Runx1", 
  	"Tec", 
  	"Notch1",
  	"Notch2",
  	"Notch3",
  	"Notch4", 
  	"Dll1", 
  	"Rbpj", 
  	"Mib1", 
  	"Myb",
    "Ets1"),
   other = c(
    "Sox9",
    "Id2",
    "Foxp2",
    "S100a4"
   )
)


write_lines(c("violin_tsne_plots\tdirectory with tsne and violin plots for injured type II derived cells, includes notch pathway genes"), 
            file.path("fig4", "README.txt"))

dir.create(file.path("fig4", "violin_tsne_plots", "notch"), recursive = T, 
             showWarnings = F)

map(notch_markers$notch, function(x){
  out_name <- file.path("fig4", 
                       "violin_tsne_plots",
                       "notch",
                       paste0(x, ".pdf"))
  
  plt <- plot_violin_tsne(atcells, 
                          .gene = x, 
                          cell_filter = naive_injured_cell_types$`Injured AT-II Derived`)
  save_plot(out_name, plt,
          base_height = 4, 
          base_aspect_ratio = 1.5)})

dir.create(file.path("fig4", 
                     "violin_tsne_plots", 
                     "tgfbeta"), 
           recursive = T, 
             showWarnings = F)

map(tgf_markers$tgfbeta, function(x){
  out_name <- file.path("fig4", 
                       "violin_tsne_plots",
                       "tgfbeta",
                       paste0(x, ".pdf"))
  
  plt <- plot_violin_tsne(atcells, 
                          .gene = x, 
                          cell_filter = naive_injured_cell_types$`Injured AT-II Derived`)
  save_plot(out_name, plt,
          base_height = 4, 
          base_aspect_ratio = 1.5)})



dir.create(file.path("fig4", 
                     "violin_tsne_plots", 
                     "other"), 
           recursive = T, 
             showWarnings = F)

map(notch_markers$other, function(x){
  out_name <- file.path("fig4", 
                       "violin_tsne_plots",
                       "other",
                       paste0(x, ".pdf"))
  
  plt <- plot_violin_tsne(atcells, 
                          .gene = x, 
                          cell_filter = naive_injured_cell_types$`Injured AT-II Derived`)
  save_plot(out_name, plt,
          base_height = 4, 
          base_aspect_ratio = 1.5)})
```


### injured naive side by side

```{r gene_sets}
new_cell_markers <- list(
  proliferation = c( 
    "Mki67", 
    "Pcna", 
    "Top2a", 
    "Tk1", 
    "Pola1"),
  at1_markers = c(
    "Pdpn",
    "Emp2",
    "Hopx",
    "Akap5",
    "Aqp5",
    "Clic5"
  ), 
  at2_markers = c(
     "Sftpc", 
     "Sftpb",
     "Lamp3", 
     "Lyz2",
     "Abca3"
  ),
  cell_cycle_arrest = c(
     "Cdkn2b", 
     "Trp53", 
     "Cdk4", 
     "Ccnd1", 
     "E2f4", 
     "E2f5"
  )
)

```

```{r by_sample_type}
dir.create(file.path("fig3", "tsne_naive_injured_by_sample_type"))
dir.create(file.path("fig3", "tsne_naive_injured_by_sample_type", 
                     "tsne_plots"))
dir.create(file.path("fig3", "tsne_naive_injured_by_sample_type", 
                     "tsne_plots", "no_labels"))

### with labels
tsne_plts <- map2(new_cell_markers,
     names(new_cell_markers),
    function(x, y){
  plts <- list()
  exp_limit = NULL
  if (x[1] == "Sftpc") {
    exp_limit = c(0, 12)
  }
  plts[[1]] <- tsne_by_sample_type(atcells, 
                    .gene = x[1],
                    cell_sets = naive_injured_cell_types,
                    top_label = names(naive_injured_cell_types),
                    limit_gene_exp = exp_limit,
                    resize_tsne = T)
  
  plts_no_label <- map(x[2:length(x)], 
           ~tsne_by_sample_type(atcells, 
                    .gene = .x,
                    cell_sets = naive_injured_cell_types, 
                    resize_tsne = T))
  plt <- plot_grid(plotlist = c(plts, plts_no_label), ncol = 1)

  save_plot(file.path("fig3", "tsne_naive_injured_by_sample_type", "tsne_plots", 
                      paste0("tsne_", y, ".pdf")),
          plt,
          ncol = 1,
          nrow = length(x),
          base_height = 2.75, 
          base_aspect_ratio = 4)
  plt
})

### without labels
map2(new_cell_markers,
     names(new_cell_markers),
    function(x, y){
  plts <- list()
  exp_limit = NULL
  if (x[1] == "Sftpc") {
    exp_limit = c(0, 12)
  }
    
  plts[[1]] <- tsne_by_sample_type(atcells, 
                    .gene = x[1],
                    cell_sets = naive_injured_cell_types,
                    limit_gene_exp = exp_limit,
                    resize_tsne = T)

  plts_no_label <- map(x[2:length(x)], 
           ~tsne_by_sample_type(atcells, 
                    .gene = .x,
                    cell_sets = naive_injured_cell_types,
                    resize_tsne = T))
       
  plt <- plot_grid(plotlist = c(plts, plts_no_label), ncol = 1)

  save_plot(file.path("fig3", "tsne_naive_injured_by_sample_type", 
                      "tsne_plots", "no_labels", 
                      paste0("tsne_", y, ".pdf")),
          plt,
          ncol = 1,
          nrow = length(x),
          base_height = 2.75, 
          base_aspect_ratio = 4)
})

```

### Cell cycle phase markers

```{r cc_phase}
cc_phase_markers <- list(
  s_phase = c( 
    "Top2a", 
    "Tk1", 
    "Pola1",
    "Ccna2",
    "Ccne1",
    "Ccne2"),
  g2_phase = c(
    "Cdk1",
    "Cdk2",
    "Ccnb1",
    "Ccnb2",
    "Ccnd2",
    "Cdkn2a",
    "Cdkn1a",
    "Cdkn1b",
    "Cdk6"
  ), 
  mito = c(
     "Mad2l1", 
     "Bub1",
     "Brca1"),
  misc = c(
     "Fgfr2", 
     "Foxm1")
)

dir.create(file.path("fig3", "tsne_injured_cell_cycle", 
                     "tsne_plots", "no_labels"), recursive = T)

write_lines("tsne_plots\ttSNE plots of cell cycle phase markers for injured type II cells (cells from injured sample with contaminating populations excluded (a small number of cells)",
            file.path("fig3", "tsne_injured_cell_cycle", "README.txt")) 
            
### with labels
tsne_plts <- map2(cc_phase_markers,
     names(cc_phase_markers),
    function(x, y){
  plts <- list()
  plts[[1]] <- tsne_by_sample_type(atcells, 
                    .gene = x[1],
                    cell_sets =  naive_injured_cell_types["Injured AT-II Derived"],
                    top_label = "Injured AT-II Derived",
                    resize_tsne = T)
  
  plts_no_label <- map(x[2:length(x)], 
           ~tsne_by_sample_type(atcells, 
                    .gene = .x,
                    cell_sets = naive_injured_cell_types["Injured AT-II Derived"],
                    resize_tsne = T))
  plt <- plot_grid(plotlist = c(plts, plts_no_label), 
                   ncol = 1)

  save_plot(file.path("fig3", "tsne_injured_cell_cycle", "tsne_plots", 
                      paste0("tsne_", y, ".pdf")),
          plt,
          ncol = 1,
          nrow = length(x),
          base_height = 2.75, 
          base_aspect_ratio = 1.33)
  plt
})


### without labels
tsne_plts <- map2(cc_phase_markers,
     names(cc_phase_markers),
    function(x, y){
  plts <- list()
  plts[[1]] <- tsne_by_sample_type(atcells, 
                    .gene = x[1],
                    cell_sets =  naive_injured_cell_types["Injured AT-II Derived"], 
                    resize_tsne = T)
  
  plts_no_label <- map(x[2:length(x)], 
           ~tsne_by_sample_type(atcells, 
                    .gene = .x,
                    cell_sets = naive_injured_cell_types["Injured AT-II Derived"],
                    resize_tsne = T))
  plt <- plot_grid(plotlist = c(plts, plts_no_label), 
                   ncol = 1)

  save_plot(file.path("fig3", "tsne_injured_cell_cycle", "tsne_plots", "no_labels",
                      paste0("tsne_", y, ".pdf")),
          plt,
          ncol = 1,
          nrow = length(x),
          base_height = 2.75, 
          base_aspect_ratio = 1.33)
  plt
})
```

## Misc

```{r dir}
dir.create("misc")
```

```{r cluster_ids}


write_lines("tsne_cluster_ids.pdf\tTSNE labeled with unbiased cluster ids",
            file.path("misc", "README.txt"))

many_cols <- c(brewer.pal(n = 12, "Paired"), 
               brewer.pal(n = 8, "Set2"), 
               brewer.pal(n = 8, "Dark2"))

atcells <- SetAllIdent(atcells, "res.1.6")
plt <- plot_tsne(atcells, ident = "res.1.6", 
                 label_text = T, label.color = "Black", 
                 .cols = many_cols) 
save_plot(file.path("misc", "tsne_cluster_ids.pdf"), 
          plt, base_height = 6, base_aspect_ratio = 1.33)

plt <- plot_tsne(atcells, ident = "res.1.6", 
                 label_text = F,  .cols = many_cols) 
save_plot(file.path("misc", "tsne_cluster_ids_no_labels.pdf"), 
          plt, base_height = 6, base_aspect_ratio = 1.33)
```

```{r clusters}
tsne_col_pal <- brewer.pal(11, "Paired")

write_lines("tsne_cell_labels.pdf\tTSNE colored by designated cell type",
            file.path("misc", "README.txt"), append = T)

set.seed(42)
new_cols <- sample(tsne_col_pal, length(tsne_col_pal))

cell_id_labels <- c("Endothelial",
  "Basal",
  "Club",
  "Ciliary",
  "Naive Type I",
  "Macrophages",
  "Injured Type II",
  "Proliferating Type II",
  "Cell Cycle Arrest Type II",
  "Transdifferentiating Type II",
  "Naive Type II")


plt <- plot_tsne(atcells, ident = "labeled_clusters", 
                 .cols = new_cols, 
                 pt.alpha = 0.75, 
                 legend_order = factor(cell_id_labels) )

save_plot(file.path("misc", "tsne_cell_labels.pdf"), 
          plt, base_height = 6, base_aspect_ratio = 1.33)
```
### GFP
```{r gfp_tsne}
write_lines(c("misc/tsne_gfp_all_cells.pdf\ttSNE of gfp expression in all cells",
              "misc/tsne_split_cells_gfp.pdf\ttSNE of gfp expression split out into cell types of interest"), 
            file.path("misc", "README.txt"), append = T)

plt <- plot_feature(atcells, gene = "gfp_counts",
             meta_data_col = T, pt.alpha = 1) +
      guides(color = guide_colorbar(barwidth = 0.5, barheight = 5)) +
      theme(axis.text = element_text(size = 10),
            axis.title = element_text(size = 10),
            legend.title = element_text(size = 10),
            legend.text = element_text(size = 10))

save_plot("misc/tsne_gfp_all_cells.pdf", plt, base_height = 6, base_aspect_ratio = 1.33)

plt <- tsne_by_sample_type(atcells, 
                    .gene = "gfp_counts",
                    cell_sets = naive_injured_cell_types,
                    top_label = names(naive_injured_cell_types), 
                    meta_data_col = TRUE)
  
save_plot(file.path("misc", paste0("tsne_split_cells_gfp.pdf")),
          plt,
          base_height = 2.75, 
          base_aspect_ratio = 4.75)

plt <- tsne_by_sample_type(atcells, 
                    .gene = "gfp_counts",
                    cell_sets = naive_injured_cell_types,
                    meta_data_col = TRUE)
  
save_plot(file.path("misc", paste0("tsne_split_cells_gfp_no_label.pdf")),
          plt,
          base_height = 2.75, 
          base_aspect_ratio = 4.75)
```
### Mito
```{r mito_tsne}

plt <- plot_feature(atcells, gene = "proportion.mito",
             meta_data_col = T, pt.alpha = 1) +
      guides(color = guide_colorbar(barwidth = 0.5, barheight = 5)) +
      theme(axis.text = element_text(size = 10),
            axis.title = element_text(size = 10),
            legend.title = element_text(size = 10),
            legend.text = element_text(size = 10))

save_plot("misc/tsne_mito_all_cells.pdf", plt, base_height = 6, base_aspect_ratio = 1.33)
```
### UMI
```{r umi_tsne}

umi_plt <- plot_feature(atcells, gene = "nUMI",
             meta_data_col = T, pt.alpha = 1) +
      guides(color = guide_colorbar(barwidth = 0.5, barheight = 5)) +
      labs(color = "# of UMIs") + 
      theme(axis.text = element_text(size = 10),
            axis.title = element_text(size = 10),
            legend.title = element_text(size = 10),
            legend.text = element_text(size = 10))

save_plot("misc/tsne_umi_all_cells.pdf", umi_plt, base_height = 6, base_aspect_ratio = 1.33)
```

### Genes
```{r genes_tsne}

gene_plt <- plot_feature(atcells, gene = "nGene",
             meta_data_col = T, pt.alpha = 1) +
      guides(color = guide_colorbar(barwidth = 0.5, barheight = 5)) +
      labs(color = "# of Genes") + 
      theme(axis.text = element_text(size = 10),
            axis.title = element_text(size = 10),
            legend.title = element_text(size = 10),
            legend.text = element_text(size = 10))

save_plot("misc/tsne_genes_all_cells.pdf", gene_plt, base_height = 6, base_aspect_ratio = 1.33)

plt <- plot_grid(umi_plt, gene_plt)
save_plot("misc/tsne_umi_genes_all_cells.pdf", plt, nrow = 1, ncol = 2,
          base_height = 3,
          base_aspect_ratio = 1.33)
```

### TSNE naive type ii and naive-non-epi
```{r}
write_lines(c("misc/tsne_naive_nonepi.pdff\ttSNE of all naive cells"), 
            file.path("misc", "README.txt"), append = T)

plt <- plot_tsne(atcells, ident = "sample_type",
          .cols = brewer.pal(3, "Set1"),
          pt.alpha = 1,
          cell_filter = naive_cell_types$all_naive) 

save_plot(file.path("misc", "tsne_naive_nonepi.pdf"), plt,
                    base_height = 6, base_aspect_ratio = 1.33)
```

### TSNE naive type ii subclusters
```{r}
write_lines(c("misc/tsne_naive_subclusters.pdf\ttSNE of naive type II subclusters (8 and 13, 11 and 12"), 
            file.path("misc", "README.txt"), append = T)

plt <- plot_tsne(atcells, ident = "sample_type",
          .cols = brewer.pal(3, "Set1"),
          pt.alpha = 1,
          cell_filter = naive_cell_types$all_naive) 

save_plot(file.path("misc", "tsne_naive_subclusters.pdf"), plt,
                    base_height = 6, base_aspect_ratio = 1.33)
```

### Just Tsne

```{r}
write_lines(c("misc/tsne_no_labels.pdf\ttSNE of all cells"), 
            file.path("misc", "README.txt"), append = T)

plt <- plot_tsne(atcells, 
          .cols = brewer.pal(3, "Set1")[2],
          pt.alpha = 1) +
  theme(legend.position = "none")

save_plot(file.path("misc", "tsne_no_labels.pdf"), plt,
                    base_height = 6, base_aspect_ratio = 1.33)
```



## Compare AT1 and AT2 profiles to "bipotent" progenitors


```{r get_dat}

bp_dat <- file.path("extdata", "nature13173-s4.txt")
if (!file.exists(bp_dat)) {  download.file("https://media.nature.com/original/nature-assets/nature/journal/v509/n7500/extref/nature13173-s4.txt", 
                                           bp_dat)
}

bp_dat <- read_tsv(bp_dat)

## exclude control cells, and non AT1 or AT2 cell types

bp_dat <- filter(bp_dat, putative_cell_type %in% c("BP", "AT1", "AT2"))
bp_metadat <- select(bp_dat, cell_name:putative_cell_type)
bp_mat <- select(bp_dat, -time_point, -sample, -putative_cell_type) %>% 
  as.data.frame() %>% 
  tibble::column_to_rownames("cell_name") %>% 
  as.matrix(.) %>% 
  t(.)


naive_markers <- read_tsv(file.path("markers",
                          "naive_cell_types_no_injured_markers_wilcox.txt")) %>% 
  filter(cluster %in% c("Naive Type I", "Naive Type II"))

#injured_markers <- read_tsv(file.path("markers",
#                          "td_injured_markers_wilcox.txt"))

top_markers <- filter(naive_markers, 
                      gene %in% rownames(bp_mat)) %>% 
  group_by(cluster) %>% 
  arrange(p_val_adj, .by_group = T) %>% 
  slice(1:25) 

expr_dat <- atcells@data[top_markers$gene, ]
bp_hmap <- bp_mat[top_markers$gene, ]

bp_metadat <- bp_metadat %>% 
  as.data.frame() %>% 
  tibble::column_to_rownames("cell_name")

bp_metadat <- bp_metadat[colnames(bp_hmap), "putative_cell_type", 
                         drop = F]

cell_samples <- unique(bp_metadat$putative_cell_type)
cols <- brewer.pal(length(cell_samples), "Paired")
names(cols) <- cell_samples

ha <- HeatmapAnnotation(bp_metadat,
                        col = list(putative_cell_type = cols))
hm <- Heatmap(bp_hmap, 
              top_annotation = ha,
              row_order = top_markers$gene,
              cluster_rows = F,
              show_row_names = F,
              show_column_names = F)
hm

```

```{r markers_from_nature}

nature_markers <- list(
  ATII = c(
    "S100g",
    "Slc34a2",
    "Lamp3",
    "Bex2",
    "Cd36",
    "Etv5",
    "Scd1",
    "Il33",
    "Fabp5",
    "Hc",
    "Sftpa1",
    "Dlk1",
    "Cxcl15",
    "Chsy1",
    "Lcn2",
    "Fasn",
    "Sftpb",
    "Egfl6",
    "Lyz1",
    "Sftpc",
    "Soat1",
    "Glrx",
    "Mlc1",
    "Retnla",
    "Mid1ip1",
    "Chi3l1",
    "Rab27a",
    "Fabp12",
    "Lgi3",
    "Lyz2"),
  ATI = c(
    "Clic5",
    "Pdpn",
    "Akap5",
    "Lmo7",
    "Ahnak",
    "Sdpr",
    "Emp2",
    "Col4a3",
    "Rtkn2",
    "Cav1",
    "Timp3",
    "Ager",
    "Msn",
    "Limch1",
    "Hopx",
    "Ptrf",
    "S100a6",
    "Dpysl2",
    "Agrn",
    "Akap2",
    "Pmp22",
    "Tinagl1",
    "Lgals3",
    "S100a14",
    "Malat1",
    "Vegfa",
    "Samhd1",
    "Qk",
    "Sema3a",
    "Aqp5")
  )

nature_markers <- bind_rows(nature_markers) %>% 
  gather(cell_type, gene) 

bp_hmap <- bp_mat[unique(nature_markers$gene), ]

marker_means <- data_frame(avg_expr = unname(rowMeans(bp_hmap)),
                           gene = names(rowMeans(bp_hmap)))
nature_markers <- left_join(nature_markers, 
          marker_means, by = "gene")

nature_markers <- nature_markers %>% 
  group_by(cell_type) %>% 
  arrange(desc(avg_expr), .by_group = T)

cell_samples <- unique(bp_metadat$putative_cell_type)
cols <- brewer.pal(length(cell_samples), "Paired")
names(cols) <- cell_samples


ha <- HeatmapAnnotation(bp_metadat,
                        col = list(putative_cell_type = cols))
hm <- Heatmap(bp_hmap, 
              top_annotation = ha,
              row_order = nature_markers$gene, 
              show_row_names = T,
              show_column_names = F,
              cluster_rows = F)
hm

```

````{r}
mdata <- atcells@meta.data[, "labeled_clusters", drop = F]
cell_samples <- unique(mdata$labeled_clusters)
cols <- brewer.pal(length(cell_samples), "Paired")
names(cols) <- cell_samples

# Chil1 == Chi3l1
nature_markers_fixed <- nature_markers
nature_markers_fixed$gene[nature_markers_fixed$gene == "Chi3l1"] <- "Chil1"
  
td_cell_types <- c(
  "Proliferating Type II",
  "Cell Cycle Arrest Type II",
  "Transdifferentiating Type II")

mdata <- mdata[mdata$labeled_clusters %in% td_cell_types, , drop = F]

ha <- HeatmapAnnotation(mdata,
                        col = list(labeled_clusters = cols))

zscores <- as.matrix(atcells@data[top_markers$gene,
                               rownames(mdata)])
hm <- Heatmap(zscores, 
              top_annotation = ha,
              row_order = top_markers$gene,
              show_row_names = F,
              show_column_names = F,
              show_column_dend = F,
              cluster_rows = F)
hm
```


```{r progression_markers}
progression_genes <- list(
  AT2_late = c(
    "Lcn2",
    "Il33",
    "Hc",
    "Trf",
    "Lyz2",
    "S100g",
    "Lyz1"),
  AT2_early = c(
    "Fabp5",
    "Lamp3",
    "Cd36",
    "Scd1",
    "Sftpb",
    "Slc34a2",
    "Abca3",
    "Sftpa1",
    "Egfl6",
    "Soat1",
    "Bex2",
    "Muc1",
    "Sftpc"
  ),
  AT1_early = c(
    "Aqp5",
    "Pdpn",
    "Rtkn2",
    "Ager",
    "Emp2",
    "Cav1",
    "Clic5",
    "Lmo7",
    "S100a6",
    "Col4a3",
    "Akap5",
    "Cryab"
  ),
  AT1_late = c(
    "Sdpr",
    "S100a14"
  )
)
```

```{r hmap, fig.height = 8}
bp_hmap <- bp_mat[unique(unlist(progression_genes)), ]

cell_samples <- unique(bp_metadat$putative_cell_type)
cols <- brewer.pal(length(cell_samples), "Paired")
names(cols) <- cell_samples
ha <- HeatmapAnnotation(bp_metadat,
                        col = list(putative_cell_type = cols))


hm <- Heatmap(bp_hmap, 
              row_order = unlist(progression_genes),
              top_annotation = ha,
              show_row_names = T,
              show_column_names = F,
              cluster_rows = F, 
              cluster_columns = T,
              row_names_side = "left")

hm@column_order
```

```{r dotplot_dev, fig.height = 12}
td_cell_types <- c(
  "Injured Type II",
  "Proliferating Type II",
  "Cell Cycle Arrest Type II",
  "Transdifferentiating Type II",
  "Naive Type I"
  )


cells_to_use <- rownames(atcells@meta.data)[atcells@meta.data$labeled_clusters %in%
                              td_cell_types]

sub_atcells <-SubsetData(atcells, cells.use = cells_to_use)
sub_atcells <- SetAllIdent(sub_atcells, "labeled_clusters")

og_plt <- DotPlot(sub_atcells, 
        genes.plot = factor(rev(unlist(progression_genes))),
        x.lab.rot = T,
        plot.legend = T,
        cols.use = brewer.pal(9, "Blues")[c(1, 9)],
        do.return = T) +
  scale_y_discrete(limits = td_cell_types) +
  coord_flip() +
  labs(colour = "Z-score\nAverage\nExpression",
       size = "Percent of cells\nwith expression")

save_plot(file.path("misc", "dotplot_development.pdf"), 
          og_plt,
          base_height = 8,
          base_aspect_ratio = 0.6)
```

```{r dotplot_nature}
## dl data from GEO
outdir <- file.path("extdata", "GSE52583")
ex_file <- file.path(outdir, 
                     "GSM1271862_E18_1_C08_IL3230.sorted.genes.fpkm_tracking.gz")
if (!file.exists(ex_file)) {
  dl_url <- "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE52583&format=file"
  
  dir.create(outdir)
  download.file(dl_url, file.path(outdir,"tmp.tar"))
  untar(file.path(outdir, "tmp.tar"), exdir = outdir)
  unlink(file.path(outdir, "tmp.tar"))
}

fpkm_files <- dir(outdir, 
                  pattern = ".gz$",
                  full.names = T)

e18_files <- str_subset(fpkm_files, "_E18_") %>% 
  .[!str_detect(., "_PTC_|_NTC")]

e18_dat <- suppressMessages(map(e18_files, read_tsv, col_names = F))
names(e18_dat) <- basename(e18_files)
e18_dat <- map(e18_dat, ~select(.x, X4, X10))
e18_dat <- bind_rows(e18_dat, .id = "library")
colnames(e18_dat) <- c("library", "gene_id", "FPKM")

e18_dat <- mutate(e18_dat,
                  cell_id = str_match(library,
                                      "E18_[0-9]_[a-zA-Z][0-9][0-9]")[, 1])

e18_dat <- dplyr::select(e18_dat, 
                         -library,
                         cell_id, gene_id, FPKM)

genes_present <- filter(e18_dat, cell_id == "E18_1_C08") %>% 
  pull(gene_id)
duplicated_genes <- unique(genes_present[duplicated(genes_present)])
e18_mat <- filter(e18_dat, 
                  !gene_id %in% duplicated_genes) %>% 
  spread(cell_id, FPKM) %>% 
  as.data.frame() %>% 
  tibble::column_to_rownames("gene_id") %>% 
  as.matrix()

## just keep BP, AT1, and AT2 cells
e18_mat <- e18_mat[, colnames(e18_mat) %in% bp_metadat$cell_name]

## make dummy seurat object with public data
dummy_obj <- atcells
dummy_obj@data <- as(log1p(e18_mat), "sparseMatrix")
mdata <- bp_metadat %>%
  mutate(orig.ident = ifelse(putative_cell_type == "AT1",
                             "Naive Type I",
                             ifelse(putative_cell_type == "AT2",
                                    "Naive Type II",
                                    "Bipotent Progenitor"))) %>% 
  as.data.frame() %>% 
  tibble::column_to_rownames("cell_name")

dummy_obj@meta.data <- mdata[colnames(e18_mat), ]
dummy_obj@cell.names <- colnames(dummy_obj@data)
dummy_obj@ident <- factor(dummy_obj@meta.data$orig.ident)

at_cell_types <- c(
  "Naive Type II",
  "Bipotent Progenitor",
  "Naive Type I"
)
nature_plt <- DotPlot(dummy_obj, 
        genes.plot = factor(rev(unlist(progression_genes))),
        x.lab.rot = T,
        plot.legend = T,
        cols.use = brewer.pal(9, "Blues")[c(1, 9)],
        do.return = T) +
  scale_y_discrete(limits = at_cell_types) +
  coord_flip() +
  labs(colour = "Z-score\nAverage\nExpression",
       size = "Percent of cells\nwith expression")

save_plot(file.path("misc", "dotplot_development_naturedata.pdf"), 
          nature_plt,
          base_height = 8,
          base_aspect_ratio = 0.6)

```


```{r plot_both}
plt <- plot_grid(og_plt, nature_plt, align = "hv")
save_plot(file.path("misc","dotplot_development_both.pdf"), 
          plt,
          nrow = 1, ncol = 2,
          base_height = 8,
          base_aspect_ratio = 0.6)
          
```


## Session Info
```{r}
sessionInfo()
```

